
---
title: "Code for Niche space of Acropora cervicornis in the Florida reef tract"
author: "Elizabeth Brown"
date: "4/7/18"
editor_options: 
  chunk_output_type: console
---
Adapted from "tutorial_cut down for running.R" Jane Elith and John Leathwick

```{r setup libraries / projections / image settings}

setwd() #set working directory to where all files are located

library(raster)
library(rgdal)
library(dismo)
library(rworldmap)

#projections
eqc<- "+proj=eqc +lat_ts=0 +lat_0=0 +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
longlat<- '+proj=longlat +ellps=WGS84 +no_defs'
utm <- "+proj=utm +zone=17 +units=km +datum=WGS84"
aea <- "+proj=aea +lat_1=24 +lat_2=31.5 +lat_0=24 +lon_0=-84 +x_0=400000 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs +towgs84=0,0,0"

#map graphics
wholeworld<-getMap(resolution="high")
FLMap<-crop(wholeworld, c(-84,-79,24,28))
FLMap<-spTransform(FLMap, longlat)

library(extrafont)
font_import()
y
loadfonts(device = "win")
```

Read in and set up biological presence/absence data 2011-2015
Training datasets - FRRP & SCREAM
Testing dataset - CREMP
```{r FRRP 2011-2015}
library(stringr)
FRRP<-read.csv("AllCorals2016B.csv")
FRRP$Year<-as.numeric(str_sub(FRRP$Date,-4,-1))
FRRPCorals2011_2015<-subset(FRRP, FRRP$Code!="L2150" & FRRP$Code!="L2151" & FRRP$Year > 2010 & FRRP$Year <2016)

FRRP_ACERV<-subset(FRRPCorals2011_2015, FRRPCorals2011_2015$Species==1)
FRRP_ACERV_nodup<- subset(FRRP_ACERV, !duplicated(FRRP_ACERV$Code))

FRRP_noACERV <- subset(FRRPCorals2011_2015, !FRRPCorals2011_2015$Code %in% FRRP_ACERV_nodup$Code)
FRRP_noACERV_nodup<- subset(FRRP_noACERV, !duplicated(FRRP_noACERV$Code))

dim(FRRP_ACERV_nodup) # 80 unique sites with ACERV presence
dim(FRRP_noACERV_nodup)# 946 unique sites with absence of ACERV
```
```{r SCREAM 2011-2015}
SCREAM<-read.csv("SCREAM_acroColDen_Site_1999_2015.csv")
Scream2011_2015<-subset(SCREAM, SCREAM$Year>2010)
Scream2011_2015$ACERV[Scream2011_2015$ACERV>0]<-1

ScreamPres<-subset(Scream2011_2015, Scream2011_2015$ACERV>0)
ScreamPres_nodup<- subset(ScreamPres, !duplicated(ScreamPres$Site))

ScreamAbs<-subset(Scream2011_2015, Scream2011_2015$ACERV==0)
ScreamAbs_nodup <- subset(ScreamAbs, !duplicated(ScreamAbs$Site))
dim(ScreamPres_nodup) # 122 unique sites with ACERV presence
dim(ScreamAbs_nodup)# 1087 unique sites with absence of ACERV

#combine FRRP and SCREAM for training dataset
PresLatitudes<-c(FRRP_ACERV_nodup$Latitude, ScreamPres$Latitude) #80+122=203 total presence points
PresLongitudes<-c(FRRP_ACERV_nodup$Longitude, ScreamPres$Longitude)
AbsLatitudes<-c(FRRP_noACERV_nodup$Latitude, ScreamAbs$Latitude)
AbsLongitudes<-c(FRRP_noACERV_nodup$Longitude, ScreamAbs$Longitude) #946+1087=2033 total absence points
length(PresLatitudes)
length(AbsLatitudes)

totalPres <- cbind(rep(1, 202), PresLatitudes, PresLongitudes) 
totalAbs <- cbind(rep(0, 2033), AbsLatitudes, AbsLongitudes)
ACERVpaFinal<-data.frame(rbind(totalPres,totalAbs))
head(ACERVpaFinal)

ACERVallxy<-data.frame(cbind( ACERVpaFinal[,3], ACERVpaFinal[,2]))
ACERVpaFinalpts<-SpatialPoints(coords=ACERVallxy, proj4string=CRS(longlat))
ACERVFinalutm<-spTransform(ACERVpaFinalpts, utm)

```
```{r CREMP 2011-2015}

CREMPlocations<-read.csv("CREMP_Locations.csv")
CREMP<-read.csv("CREMP_Core_Species_Richness_Table.csv")
CREMPall<-merge(CREMP, CREMPlocations, by="siteid")
CREMPall2011<-CREMPall[(CREMPall$Year>=2011)&(CREMPall$Year<=2015),]

CREMPpres<-subset(CREMPall2011, CREMPall2011$Acropora_cervicornis==1)
CREMPpresNoDup<-subset(CREMPpres, !duplicated(CREMPpres$siteid))

CREMPabs <- subset(CREMPall2011, !CREMPall2011$siteid %in% CREMPpresNoDup$siteid)
CREMPabsNoDup<- subset(CREMPabs, !duplicated(CREMPabs$siteid))
head(CREMPabs)
CREMPunique<-rbind(CREMPpresNoDup, CREMPabsNoDup)

testCREMP<-SpatialPoints(cbind(CREMPunique$lonDD, CREMPunique$latDD))

```

Figure 2. Presence locations used to train and test the Acropora cervicornis niche model along the Florida reef tract (2011-2015).
```{r Site figure }
library(ggmap)

#acropora critical habitat shapefile
critical <- readOGR(".","Acropora_Critical_Habitat")

#Get map
m <- get_map(c(-81.5,25.6),zoom=8,maptype="satellite",source="google")
ggmap(m) + 
  geom_polygon(data = fortify(critical),
                  aes(long, lat, group = group),
                  fill = "black", lwd=1, alpha = 0) +
  geom_point(data = ScreamPres_nodup,
             aes(x = Longitude, y = Latitude), shape = 3,cex=3,
             color = "white", alpha = 0.7) + 
  geom_point(data = FRRP_ACERV_nodup,
             aes(x = Longitude, y = Latitude), shape = 3,cex=3,
             color = "red", alpha = 0.7) + 
  geom_point(data = CREMPpresNoDup,
             aes(x = lonDD, y = latDD), shape = 3,cex=3,
             color = "yellow", alpha = 0.7)+

   theme(legend.justification = c(0, 0), legend.position = c(0, 1))
ggsave("Fig2back.png", height=6.5, units = "in")

#inset map for figure 2
plot_Fig1_inset <- function (){
  plot(wholeworld, xlim=c(-130,-35), ylim=c(-10,40), col = "black", xaxt="n", yaxt="n")
  plot(extent(c(-83,-80,24.3,27.3)), add= TRUE, col = "green3", lwd = 3)
    maps::map.scale(x= -125, y= -10, relwidth=0.3, ratio=FALSE)
    axis(1, at=seq(-120,-30.,20),seq(-120,-30.,20), NA, cex.axis=1, font=1, tck=.02)
    axis(2, at=seq(-20,40,20), seq(-20,40,20), cex.axis=1, font=1, tck=.02)
    axis(3, at=seq(-120,-30.,20),seq(-120,-30.,20), NA, cex.axis=1, font=1, tck=.02)
    axis(4, at=seq(-20,40,20), seq(-20,40,20), cex.axis=1, font=1, tck=.02)
    text(c(-120,-80), -25, c("120°W", "80°W"))
    text( -126, c(0,20), c("0", "20°N"))
    box(which="plot")
    arrows(-52, 25, -52, 37, length=0.1)
    text(-52, 30, "N")
}
png( "Fig2inset.png",  width= 3,  height= 3.1,  units = "in",
  res   = 300,  pointsize = 8)
plot_Fig1_inset() #print figure
dev.off()
```

Read in environmental raster data
Output files needed for the BRT
```{r Environmental variables}

#Create blank reef mask from Unified Florida Reef Tract polygon
unified <- readOGR(".","UnifiedFloridaReefTract_poly")
unifiedutm<- spTransform(unified, utm)

criticalutm<-spTransform(critical, utm)
criticalutmcrop<-crop(criticalutm, extent(288.9702, 597.0663, 2702.092, 2936.424))
northunifiedutmcrop<-crop(unifiedutm, extent(288.9702, 597.0663, 2936.424, 3016.051))
northunifiedcropaggregated<-aggregate(northunifiedutmcrop)
criticalutmcropAggregated<-aggregate(criticalutmcrop)

allReef<-bind( criticalutmcrop,northunifiedcropaggregated)

emptyRaster <- raster(resolution = c(1, 1), xmn = 298.5721, xmx = 601.0686, ymn= 2700.835, ymx = 3016.051, crs = utm)
emptyRaster[]<-1
extentmask<-mask(emptyRaster, unifiedutm)

#(i) depth, (ii & iii) mean and variance of daily sea surface temperature, (iv) chlorophyll-a concentrations, (v) turbidity, (vi) wave energy, and (vii) distance from coast

#(i) depth
depthm <- "depthm.txt"
depthmr<- raster(depthm, values=TRUE)
e=extent(481304.821426, 807304.821426, 32335.4432404, 377335.44324)
extentraster<-raster(e, nrows=326, ncols=345, crs=aea)
proj4string(depthmr)<-aea
Depthmutm<-projectRaster(depthmr,crs=utm)
depth_1km<-resample(Depthmutm, extentmask, method="ngb")

#(ii) mean of daily sea surface temperature
meanDailySST<-raster("meanDailySST2011-2015.tif")
meanDailySSTutm<- projectRaster(meanDailySST, crs=utm)
meanDailySST_1km<-resample(meanDailySSTutm, extentmask, method="bilinear")
#plot(meanDailySST_1km)

#(iii) variance of daily sea surface temperature
varDailySST<-raster("varianceDailySST2011-2015.tif")
varDailySSTutm<- projectRaster(varDailySST, crs=utm)
varDailySST_1km<-resample(varDailySSTutm, extentmask, method="bilinear")
#plot(varDailySST_1km)

#(iv) chlorophyll-a concentrations
meanChla2011_2015<-raster("meanDailyChla2011-2015_1km.asc")
proj4string(meanChla2011_2015)<-eqc
meanChla2011_2015utm<-projectRaster(meanChla2011_2015, crs=utm)
meanChla2011_2015_1km<-resample(meanChla2011_2015utm, extentmask, method="bilinear")
#plot(meanChla2011_2015_1km)

#(v) turbidity
meanDailyKD490<-raster("avgDailyKD4902013-2015.tif")
meanDailyKD490utm<-projectRaster(meanDailyKD490, crs=utm)
kd490_1km<-resample(meanDailyKD490utm, extentmask, method = "bilinear")
kd490_1km<-mask(kd490_1km, extentmask)
#plot(kd490_1km)

#(vi) wave energy
waveEnergy<- raster("AverageDailyWE2011-2015.tif")
waveEnergyutm<-projectRaster(waveEnergy, crs=utm)
waveEnergyutmkJ<-waveEnergyutm/1000
waveEnergy_1km<-resample(waveEnergyutmkJ, extentmask, method="bilinear")
#plot(waveEnergy_1km)

#(vii) distance from coast
land_utm<-spTransform(FLMap, utm)
# Rasterize land pixels
r2 <- rasterize(land_utm, extentmask, 1)
# Calculate distance to nearest non-NA pixel
d <- distance(r2)
distFromCoast<-mask(d, extentmask)
#plot(distFromCoast)

#####output asciis

writeRaster(extentmask,'blank_1km.asc', format='ascii', overwrite=TRUE)
writeRaster(depth_1km,'depth_1km.asc', format='ascii', overwrite=TRUE)
writeRaster(meanDailySST_1km,'meanDailySST_2011_2015.asc', format='ascii', overwrite=TRUE)
writeRaster(varDailySST_1km,'varDailySST_2011_2015.asc', format='ascii', overwrite=TRUE)
writeRaster(meanChla2011_2015_1km,'meanChla_2011_2015.asc', format='ascii', overwrite=TRUE)
writeRaster(kd490_1km,'kd490_2013_2015.asc', format='ascii', overwrite=TRUE)
writeRaster(waveEnergy_1km,'WE_2011_2015.asc', format='ascii', overwrite=TRUE)
writeRaster(distFromCoast,'distFromCoast.asc', format='ascii', overwrite=TRUE)

allPredictors<-stack(depth_1km, meanDailySST_1km, varDailySST_1km, meanChla2011_2015_1km,  kd490_1km, waveEnergy_1km, distFromCoast)
names(allPredictors) <- c( "Depth","MeanSST", "VarianceSST","Chlorophylla",  "Turbidity","WaveEnergy", "distFromCoast")

finalPredictors<-stack(depth_1km, meanDailySST_1km, varDailySST_1km,   kd490_1km, waveEnergy_1km, distFromCoast)
names(finalPredictors) <- c( "Depth","MeanSST", "VarianceSST","Turbidity","WaveEnergy", "distFromCoast")

sdmdataFinal<- extract(finalPredictors, ACERVFinalutm)
ACERVpaDataFinal<- cbind(ACERVpaFinal,sdmdataFinal)
names(ACERVpaDataFinal)<- c("pa", "latitude", "longitude", "Depth", "MeanSST", "VarianceSST",  "Turbidity", "WaveEnergy", "DistanceFromCoast")
write.csv(ACERVpaDataFinal, "ACERV_pa_data.csv")
```

S1 Figure. Environmental raster data.
```{r Environmental raster data. }
library(GISTools)
# create function to add Florida map, axes, scale bar, and north arrow

FLMapfun <- function() {
    plot(FLMap, add = TRUE, col = "gray")
    box(which="plot", lwd=1)
    axis(1, at=seq(-83,-80.,1), NA, cex.axis=3, tck=.02)
    axis(2, at=seq(25,27.,1), NA, cex.axis=3, tck=.02)
    axis(3, at=seq(-83,-80.,1), NA, cex.axis=3, tck=.02)
    axis(4, at=seq(25,27.,1), NA, cex.axis=3, tck=.02)
    maps::map.scale(x= -82.9, y= 25.7, relwidth=0.2, ratio=FALSE)
    arrows(-83, 26.5, -83, 27, length=0.1)
    text(-83, 26.7, "N", cex=1.2)
    text(c(-81,-80.1), 24.5, c("81°W", "80°W"))
    text(-82.9, seq(25,26,1), c("25°N", "26°N"))
}

predictors_longlat<-projectRaster(allPredictors, crs = longlat)

#function to plot variables and add map elements
plot_S1_Figure <- function(){
  plot(predictors_longlat, 
       main = c("Depth (m)",  "Mean Daily SST (°C)", expression(paste("Thermal Variability (",degree,"C)")), expression(paste("Chlorophyll-a (mg ", m^{-3})~ ")"), expression(paste("Turbidity (", m^{-1})~ ")"), expression(paste("Wave Energy (kJ ", m^{-2})~ ")"), "Distance From Coast (km)"),
       yaxt='n', xaxt='n', 
       legend.width=1,
       addfun = FLMapfun, nc = 3, nr = 3, family = "Arial")

}
png(  "S1Figure.png",  width     = 5.5,  height    = 4.5,  units= "in",
  res       = 600,  pointsize = 8
)
plot_S1_Figure() #print figure
dev.off()

```

S2 Figure. Correlogram of environmental predictors.
```{r  Correlogram of environmental predictors }
library(corrplot)

par(mar=c(2,2,2,2))
predictor_correlations=layerStats(allPredictors, 'pearson', na.rm=T)
corr_matrix=predictor_correlations$'pearson correlation coefficient'

corrplot(corr_matrix, method="color", addCoef.col = "black",sig.level = 0.05)
```

Fit Boosted Regression tree 
-The code to run the model has been commented out because every iteration introduces some stochasticity. To run my exact model, read in "ACERV_BRT_Model.rds"
```{r fit model acerv.brt }

source("brt.functions.R")
library(gbm)
# set starting number of trees to 30
# Set tree complexity to 5
# set learning rate to 0.005
# set bag fraction to 0.75

#  set.seed(0)
#ACERV_BRT_Model <- gbm.step(
#  data = ACERVpaDataFinal,
#  gbm.x = 4:9, #columns of predictor variables
#  gbm.y = 1, #column of pa response variable
#  family = "bernoulli",
#  n.trees = 30,
#  tree.complexity = 5,
#  learning.rate = 0.005,
#  bag.fraction = 0.75
#)
#save model
#saveRDS(ACERV_BRT_Model, file = "ACERV_BRT_Model.rds", ascii = FALSE)
#load model

ACERV_BRT_Model<-readRDS("ACERV_BRT_Model.rds", refhook = NULL)

```

Model statistics
```{r model evaluation}
ACERV_BRT_Model$n.trees
ACERV_BRT_Model$bag.fraction
ACERV_BRT_Model$self.statistics
ACERV_BRT_Model$cv.statistics

mean(ACERV_BRT_Model$cv.roc.matrix)
sd(ACERV_BRT_Model$cv.roc.matrix)
```

Output the SDM
-Input a "blank" ascii of 1s as the prediction space and 0s as noData
-Input individual asciis of each predictor
```{r predict SDM to acerv_preds.asc}
#Check the order of the column in the presence/absence data
names(ACERVpaDataFinal)[4:9]

#Make grid names the same order with the prediction space first
keys.grid.names <-
  c(
    "blank_1km.asc",  "depth_1km.asc", "meanDailySST_2011_2015.asc",  "varDailySST_2011_2015.asc", "kd490_2013_2015.asc","newWE_1km.asc", "distFromCoast.asc"  )

#store the predictor names and assign each ascii to a list
keys.variable.names <-
  c("blank", names(ACERVpaDataFinal)[4:9]) # here make sure the order is the same as above, if you're using different data

for (i in 1:length(keys.grid.names)) {
  assign(keys.variable.names[i],
         scan(keys.grid.names[i], skip = 6, na.string = "-9999"),
         pos = 1)
}
# predict to a dataframe
keys.pred <-
  data.frame (
    Depth, MeanSST, VarianceSST, Turbidity, WaveEnergy, DistanceFromCoast )
#names(keys.pred)<- c("Depth", "MeanSST", "VarianceSST", "Turbidity", "WaveEnergy","DistanceFromCoast")

# remove first row
keys.pred <- keys.pred[2:95130,]
keys.pred <- keys.pred[!is.na(blank), ]

par(mfrow = c(1,1))
par(mar = c(2,2,2,2))
#gbm.predict.grids(
#  ACERV_BRT_Model,
#  keys.pred,
#  want.grids = T,
#  sp.name = "ACERV_prediction", #output prediction raster
#  pred.vec = rep(-9999, 95130),
#  filepath = "",
#  num.col = 302,
#  num.row = 315,
#  xll = 0,
#  yll = 0,
#  cell.size = 1,
#  no.data = -9999,
#  plot = T #plots the SDM
#)

```

Read in acerv_preds.asc and plot SDMs
Figure 4. Boosted regression tree or ‘niche’ model that predicted the probability of occurrence of Acropora cervicornis along the Florida reef tract (2011 – 2015). 
```{r Florida reef tract prediction map}

library(sp)
library(maps)
library(GISTools)

blank<-raster("blank_1km.asc")
acerv_map<-raster("ACERV_prediction.asc")
extent(acerv_map) <- extent(blank)
proj4string(acerv_map) <-utm
longlatMap<- projectRaster(acerv_map, crs = longlat)

critical_longlat<- spTransform(allReef, longlat)
criticalMap<-mask(longlatMap, critical_longlat)

threshBoundary<-criticalMap
threshBoundary[threshBoundary<.09]<-NA
threshBoundary09 <- rasterToPolygons(threshBoundary>.09, dissolve=TRUE)

threshBoundary<-criticalMap
threshBoundary[threshBoundary<.27]<-NA
threshBoundary27 <- rasterToPolygons(threshBoundary>.27, dissolve=TRUE)

breaks<-seq(0,0.6,by=0.01)
colors0.6<-rev(terrain.colors(60))
probOccur<-list(text='Probability of Occurrence', 
                side=4, line=3)
scale0.6<- list(at=c(0,0.1,0.2,0.3,0.4,0.5,0.6),
                        labels=c(0,0.1,0.2,0.3,0.4,0.5,0.6))
par(mar=c(1,1,1,1))
#Florida reef tract
plot_Fig4 <- function() {
    par(oma=c(1, 1, 1, 1), mar=c(0,0,0,1))
    #dev.new(width=6, height=8)
    plot(criticalMap, ylim=c(24.35972, 27.32156), legend.args=probOccur, axis.args= scale0.6, xaxt="n", yaxt="n",
         tck = 0.02, col = colors0.6, breaks = breaks)
    plot(FLMap, add=TRUE, col = "grey")
    plot(threshBoundary09, border = "blue",lty=2, add=T)
    plot(threshBoundary27, border = "red", lty=3, add=T)
    axis(1, at=seq(-83,-80,1), NA, cex.axis=.7, font=1, tck=.02)
    axis(2, at=seq(25,27,1), NA, cex.axis=.7, font=1, tck=.02)
    axis(3, at=seq(-83,-80,1), NA, cex.axis=.7, font=1, tck=.02)
    axis(4, at=seq(25,27,1), NA, cex.axis=.7, font=1, tck=.02)
    text(c(-81, -80.1), 24.5, c("81°W", "80°W"))
    text(-82.9, seq(25.0,26.0,1), c("25°N", "26°N"))
    maps::map.scale(x= -82.7, y= 25.5, relwidth=0.2, ratio=FALSE)
    arrows(-82.8, 26.4, -82.8, 26.65, length=0.1)
    text(-82.8, 26.5, "N")
    box(which = "plot")
    legend("topleft",  col=c("blue","red"),lty=c(2,3), title="Threshold", c("> 9%", "> 27%"), pt.cex=2)
}

png(  "Fig4.png",  width     = 4.5,  height= 4,  units = "in",  res = 600,  pointsize = 8)
plot_Fig4() #print figure
dev.off()

```

S4 Figure. Probability of occurrence of Acropora cervicornis in the Florida reef tract (2011 – 2015) above the 9% and 27% threshold values subregional detail maps. 
(a) northern reef tract (b) Biscayne (c) Upper Keys (d) Middle Keys, (e) lower Keys,  (f) Marquesas, and (g) Dry Tortugas Subregional detail maps
```{r Subregional detail maps}
library(rgdal)
library(maptools)
subregions <- readOGR(".","Unified_Florida_Reef_Map_v2.0_Regions")
northReef<- subset(subregions, OBJECTID %in% c(2,3,4,5,12))
northAggregated<-aggregate(northReef)
mydata=c("1","2","3")
names(mydata)<-c("Region", "SHAPE_Leng", "SHAPE_Area")

southReefAll<-subset(subregions, OBJECTID %in% c(1,7,8,9,10,11))
subregions<-bind(southReefAll,northAggregated)
subregionslonglat<- spTransform(subregions, longlat)
subregionsutm<- spTransform(subregions, utm)

northernReef<- extent(subregionslonglat[7,])
biscayne<- extent(subregionslonglat[2,])
upperKeys<- extent(subregionslonglat[6,])
middleKeys<- extent(subregionslonglat[3,])
lowerKeys<- extent(subregionslonglat[5,])
marquesas<- extent(subregionslonglat[1,])
dryTortugas<-extent(subregionslonglat[4,])

#h) key
plot_S4Fig_h <- function(){
  par(oma=c(0, 0, 0, 1), mar=c(0,0,0,0))
  plot(FLMap, xlim=c(-83.15, -79.93873), ylim=c(24.3, 27.2643), col = "gray")
  plot(subregionslonglat, add=T, col="blue")
  plot(FLMap,col = "gray30", add=T)
  box(which="plot")
  axis(1, at=seq(-83,-80,1), NA, cex.axis=.7, font=1, tck=.02)
  axis(2, at=seq(25,27,1), NA, cex.axis=.7, font=1, tck=.02)
  axis(3, at=seq(-83,-80,1), NA, cex.axis=.7, font=1, tck=.02)
  axis(4, at=seq(25,27,1), NA, cex.axis=.7, font=1, tck=.02)
  text(c(-81, -80.15), 24.3, c("81°W", "80°W"))
  text(-82.9, seq(25.0,26.0,1), c("25°N", "26°N"))
  maps::map.scale(x= -82.7, y= 25.5, relwidth=0.2, ratio=FALSE)
  arrows(-83, 26.8, -83, 27.3, length=0.1)
  text(-83, 27, "N")
  text(subregionslonglat, c("b","f","d","a","c","e","g",""), cex=1.5, col = "white")
}

png( "S4h.png",  width= 2,  height= 2,  units = "in",
  res   = 600,  pointsize = 9)
plot_S4Fig_h() #print figure
dev.off()

#a) Dry Tortugas-83.3,-82.6)24.5,24.9
plot_S4Fig_a <- function(){
    par(oma=c(0, 0, 0, 1), mar=c(0,0,0,1))
    plot(criticalMap,
    xlim = c(dryTortugas@xmin,dryTortugas@xmax) , ylim=c(dryTortugas@ymin, dryTortugas@ymax),
    legend.args=probOccur, axis.args= scale0.6, 
    xaxt="n",yaxt="n", col = colors0.6, breaks = breaks)
    plot(threshBoundary09, border = "blue", lty=2, add=T)
    plot(threshBoundary27, border = "red", lty=3, add=T)
    
    axis(1, at=seq(-83,-81.8,.1), NA, font=1, tck=.02)
    axis(2, at=seq(24.4,25,.1), NA, font=1, tck=.02)
    axis(3, at=seq(-83,-81.8,.1), NA, font=1, tck=.02)
    axis(4, at=seq(24.4,25,.1), NA, font=1, tck=.02)
    
    text(c(-82.8,-82.64), 24.75, c("82.8°W","82.6°W"))
    text(-83.03, seq(24.6,24.7,.1), c("24.6°N","24.7°N"))
    
    legend("bottomleft", pch = c(20,NA,NA), lty=c(NA,2,3), col=c("grey","blue","red"), c("Land", "> 9% Threshold","> 27% Threshold"), pt.cex=2)
    
    maps::map.scale(x= -82.79, y= 24.57, relwidth=0.2, ratio=FALSE)
    arrows(-82.68, 24.66, -82.68, 24.72, length=0.1)
    text(-82.68, 24.68, "N")
    
}
png( "S4a.png",  width= 3,  height= 2,  units = "in",
  res   = 600,  pointsize = 9)
plot_S4Fig_a() #print figure
dev.off()

#b) marquesas
plot_S4Fig_b <- function(){
    par(oma=c(0, 0, 0, 1), mar=c(0,0,0,1))
    plot(criticalMap,
         xlim = c(marquesas@xmin, marquesas@xmax),
         ylim=c(24.4,24.7), 
         legend.args=probOccur, axis.args= scale0.6, 
         xaxt='n',yaxt='n', col = colors0.6, breaks = breaks)
    plot(FLMap, add=TRUE, col = "grey")
    plot(threshBoundary09, border = "blue", lty=2, add=T)
    plot(threshBoundary27, border = "red", lty=3, add=T)
    
    axis(1, at=seq(-81.8,-82.8,-.1), NA, font=1, tck=.02)
    axis(2, at=seq(24.4,24.7,.1), NA, font=1, tck=.02)
    axis(3, at=seq(-81.8,-82.8,-.1), NA, font=1, tck=.02)
    axis(4, at=seq(24.4,24.7,.1), NA, font=1, tck=.02)
    text(seq(-82.2,-82,.2), 24.35, c("82.2°W","82°W"))
    text(-82.52, seq(24.5,24.6,.1), c("24.5°N","24.6°N"))
    
    legend("topleft", pch = c(20,NA,NA), lty=c(NA,2,3), col=c("grey","blue","red"), c("Land", "> 9% Threshold","> 27% Threshold"), pt.cex=2)
    maps::map.scale(x= -82.4, y= 24.5, relwidth=0.2, ratio=FALSE)
    arrows(-81.9, 24.67, -81.9, 24.74, length=0.1)
    text(-81.9, 24.7, "N")
    box(which = "plot")
}
png( "S4b.png",  width= 4,  height= 2,  units = "in",
  res   = 600,  pointsize = 10)
plot_S4Fig_b() #print figure
dev.off()

#c) Lower keys
plot_S4Fig_c <- function(){
      par(oma=c(0, 0, 0, 1), mar=c(0,0,0,1))
      plot(criticalMap,
           xlim = c(lowerKeys@xmin, lowerKeys@xmax),
           ylim=c(lowerKeys@ymin,24.8), 
           legend.args=probOccur, axis.args= scale0.6, xaxt='n',yaxt='n',
            col = colors0.6, breaks = breaks)
      plot(FLMap, add=TRUE, col = "grey")
      plot(threshBoundary09, border = "blue", lty=2, add=T)
      plot(threshBoundary27, border = "red", lty=3, add=T)
      axis(1, at=seq(-82.6,-81.2,.1), NA, font=1, tck=.02)
      axis(2, at=seq(24.4,24.7,.1), NA, font=1, tck=.02)
      axis(3, at=seq(-82.6,-81.2,.1), NA, font=1, tck=.02)
      axis(4, at=seq(24.4,24.7,.1), NA, font=1, tck=.02)
      
      text(c(-81.4,-81.23), 24.4, c("81.4°W","81.2°W"))
      text(-81.75, seq(24.4,24.6,.2), c("24.4°N","24.6°N"))
      
      legend("topleft", pch = c(20,NA,NA), lty=c(NA,2,3), col=c("grey","blue","red"), c("Land", "> 9% Threshold","> 27% Threshold"), pt.cex=2)
      maps::map.scale(x= -81.5, y= 24.5, relwidth=0.3, ratio=FALSE)
      arrows(-81.25, 24.74, -81.25, 24.81, length=0.1)
      text(-81.25, 24.77, "N")
}

png( "S4c.png",  width= 3.2,  height= 2,  units = "in",
  res   = 600,  pointsize = 9)
plot_S4Fig_c() #print figure
dev.off()

#d) middle keys
plot_S4Fig_d <- function(){
    par(oma=c(0, 0, 0, 1), mar=c(0, 0, 0, 1))
    plot(criticalMap, xlim = c(middleKeys@xmin, middleKeys@xmax),      ylim=c(middleKeys@ymin,middleKeys@ymax), 
         legend.args=probOccur, axis.args= scale0.6, 
         xaxt='n',yaxt='n', col = colors0.6, breaks = breaks)
    plot(threshBoundary09, border = "blue", lty=2, add=T)
    plot(threshBoundary27, border = "red", lty=3, add=T)
    plot(FLMap, add=TRUE, col = "grey")
    
    axis(1, at=seq(-81.2,-80.6,.2), NA, font=1, tck=.02)
    axis(2, at=seq(24.6,24.9,.1), NA, font=1, tck=.02)
    axis(3, at=seq(-81.2,-80.6,.2), NA, font=1, tck=.02)
    axis(4, at=seq(24.6,24.9,.1), NA, font=1, tck=.02)
    
    text(c(-81.,-80.6), 24.53, c("81°W","80.6°W"))
    text(-81.17, c(24.6,24.8), c("24.6°N","24.8°N"))
    
    legend("topleft", pch = c(20,NA,NA), lty=c(NA,2,3), col=c("grey","blue","red"), c("Land", "> 9% Threshold","> 27% Threshold"), pt.cex=2)
    maps::map.scale(x= -80.9, y= 24.66, relwidth=0.2, ratio=FALSE)
    arrows(-80.75, 24.92, -80.75, 24.99, length=0.1)
    text(-80.75, 24.95, "N")
    box(which = "plot")
}

png( "S4d.png",  width= 3.2,  height= 2,  units = "in",
  res   = 600,  pointsize = 9)
plot_S4Fig_d() #print figure
dev.off()

#e) upper keys
plot_S4Fig_e <- function(){
    par(oma=c(0, 0, 0, 1), mar=c(0,0,0,1))
    plot(criticalMap, xlim = c(upperKeys@xmin, upperKeys@xmax), ylim=c(upperKeys@ymin,upperKeys@ymax),  legend.args=probOccur, axis.args= scale0.6, 
         tck = 0.02, col = colors0.6, breaks = breaks, xaxt='n',yaxt='n')
    plot(FLMap, add=TRUE, col = "grey")
    plot(threshBoundary09, border = "blue", lty=2, add=T)
    plot(threshBoundary27, border = "red", lty=3, add=T)
   
    axis(1, at=seq(-80.6,-80.2,.1), NA, cex.axis=.7, font=1, tck=.02)
    axis(2, at=seq(24.9,25.3,.1), NA, cex.axis=.7, font=1, tck=.02)
    axis(3, at=seq(-80.6,-80.2,.1), NA, cex.axis=.7, font=1, tck=.02)
    axis(4, at=seq(24.9,25.3,.1), NA, cex.axis=.7, font=1, tck=.02)
    
    text(seq(-80.4,-80.2,.2), 24.86, c("80.4°W", "80.2°W"))
    text(-80.58, seq(25,25.2,.2), c("25°N","25.2°N"))
    legend("topleft", pch = c(20,NA,NA), lty=c(NA,2,3), col=c("grey","blue","red"), c("Land", "> 9% Threshold","> 27% Threshold"), pt.cex=2)
    maps::map.scale(x= -80.4, y= 24.93, relwidth=0.3, ratio=FALSE)
    arrows(-80.2, 25.08, -80.2, 25.13, length=0.1)
    text(-80.2, 25.1, "N")
    box(which = "plot")
}

png( "S4e.png",  width= 3.4,  height= 3,  units = "in",
  res   = 600,  pointsize = 9)
plot_S4Fig_e() #print figure
dev.off()

#f) biscayne
plot_S4Fig_f <- function(){
    par(oma=c(0, 0, 0, 1), mar=c(0,0,0,1))
    plot(criticalMap, xlim = c(biscayne@xmin, -80.04),
         ylim=c(biscayne@ymin,biscayne@ymax),  legend.args=probOccur, axis.args= scale0.6, 
         tck = 0.02, col = colors0.6, breaks = breaks, xaxt='n',yaxt="n")
    plot(threshBoundary09, border = "blue", lty=2, add=T)
    plot(threshBoundary27, border = "red", lty=3, add=T)
    plot(FLMap, add=TRUE, col = "grey")
    
    axis(1, at=seq(-80.3,-80.,.1), NA, cex.axis=.7, font=1, tck=.02)
    axis(2, at=seq(25.3,25.7,.1), NA, cex.axis=.7, font=1, tck=.02)
    axis(3, at=seq(-80.3,-80.,.1), NA, cex.axis=.7, font=1, tck=.02)
    axis(4, at=seq(25.3,25.7,.1), NA, cex.axis=.7, font=1, tck=.02)
    
    text(c(-80.2,-80.1), 25.28, c("80.2°W","80.1°W"))
    text(-80.35, seq(25.3,25.5,.2), c("25.3°N","25.5°N"))
    
    legend("topleft", pch = c(20,NA,NA), lty=c(NA,2,3), col=c("grey","blue","red"), c("Land", "> 9% Threshold","> 27% Threshold"), pt.cex=2)
    maps::map.scale(x= -80.29, y= 25.6, relwidth=0.2, ratio=FALSE)
    arrows(-80.07, 25.31, -80.07, 25.37, length=0.1)
    text(-80.07, 25.34, "N")
    box(which = "plot")
}

png( "S4f.png",  width= 3.1,  height= 3,  units = "in",
  res   = 600,  pointsize = 9)
plot_S4Fig_f() #print figure
dev.off()

#g) Northern reef tract
dev.off()
plot_S4Fig_g <- function(){
    par(oma=c(0, 0, 0, 1), mar=c(0,0,0,1))
    plot(criticalMap, xlim = c(northernReef@xmin, northernReef@xmax), family = "Arial", ylim=c(northernReef@ymin, northernReef@ymax), legend.args=probOccur, axis.args= scale0.6, 
        xaxt='n', yaxt='n', col = colors0.6, breaks = breaks)
    plot(threshBoundary09, border = "blue", lty=2, add=T)
    plot(threshBoundary27, border = "red", lty=3, add=T)
    plot(FLMap, add=TRUE, col = "grey")
    text(c(-80.2,-80), 25.8, c("80.2°W","80°W"))
    text(-80.33, c(26,27), c("26°N","27°N") )
    
    axis(1, at=seq(-80,-80.2,-.2), NA, cex.axis=.7, font=1, tck=.02)
    axis(2, at=seq(25,27,.2), NA, cex.axis=.7, font=1, tck=.02)
    axis(3, at=seq(-80,-80.2,-.2), NA, cex.axis=.7, font=1, tck=.02)
    axis(4, at=seq(25,27,.2), NA, cex.axis=.7, font=1, tck=.02)
    legend("topleft", pch = c(20,NA,NA), lty=c(NA,2,3), col=c("grey","blue","red"), c("Land", "> 9% Threshold","> 27% Threshold"), pt.cex=2)

    maps::map.scale(x= -80.35, y= 26.7, relwidth=0.3, ratio=FALSE)
    arrows(-80, 25.87, -80, 25.93, length=0.1)
    text(-80, 25.9, "N")
    box(which = "plot")
}

png( "S4g.png",  width= 2.5,  height= 5.25,  units = "in",
  res   = 600,  pointsize = 9)
plot_S4Fig_g() #print figure
dev.off()

```

Using package spatialEco, calculate amount of area above different thresholds.
```{r zonal statistics}
library(stats)
library(spatialEco)
library(raster)

# area by subregion
threshold0.09 <- function(x, t=0.09) {
  if ( length(x[x >= t]) < 1 )  return(0) 
    if ( length(x[x >= t]) == length(x) ) return(1) 
     else return( length(x[x >= t]) ) 
}
threshold0.27 <- function(x, t=0.27) {
  if ( length(x[x >= t]) < 1 )  return(0) 
    if ( length(x[x >= t]) == length(x) ) return(1) 
     else return( length(x[x >= t]) ) 
}
subregions0.09 <- zonal.stats(x=subregionslonglat, y=criticalMap, stat=threshold0.09, trace=TRUE, plot=TRUE)
par(mar=c(0,0,0,0))
plot(criticalMap>.09)
plot(subregionslonglat, add=T)
subregions0.27 <- zonal.stats(x=subregionslonglat, y=criticalMap, stat=threshold0.27, trace=TRUE, plot=TRUE)

thresholdList<-cbind(as.character(subregionslonglat$Region), subregions0.09,subregions0.27)

subregionsums<- c("Total", sum(as.numeric(thresholdList[,2]), na.rm = T), sum(as.numeric(thresholdList[,3]), na.rm = T))

thresholdListComplete<-rbind(thresholdList, subregionsums)
thresholdListComplete
```

Figure 3. Relative importance of environmental predictors
```{r Partial dependency plots of the boosted regression tree model}

#plot fitted vs predictors and relative importance
gbm.plot(ACERV_BRT_Model, n.plots = 6, plot.layout = c(2,3), smooth = TRUE, write.title = F, cex.axis = 2, cex.lab = 2,  rug.tick = 0.04)
 
```

S3 Figure. Interaction plots for the boosted regression tree niche model.
```{r interactions}

acerv.brt.int <- gbm.interactions(ACERV_BRT_Model)
acerv.brt.int$rank.list
acerv.brt.int$interactions

#ACERV_BRT_Model$var.names<- c("Depth", "MeanSST", "Variance of", "Turbidity", "WaveEnergy","DistanceFromCoast")

png( "S3A.png",  width= 5,  height= 5,  units = "in",
  res   = 300,  pointsize = 8)
par(mfrow=c(1,1))
par(mar=c(1,1,0,0))

gbm.perspec(ACERV_BRT_Model, x.label = "Wave Energy", y.label = "Turbidity(K490)", 
            5,
            4,
            x.range = c(0, 2.5),
            y.range = c(0, 0.3),
            z.range = c(0, .3))
dev.off()

png( "S3B.png",  width= 5,  height= 5,  units = "in",
  res   = 300,  pointsize = 8)
gbm.perspec(ACERV_BRT_Model,x.label = "Mean Daily SST", y.label = "Thermal Variability",
            2,
            3,
            x.range = c(20,30),
            y.range = c(0,50 ),
            z.range = c(0, 0.7))
dev.off()
```

Confusion matrices
S5 Figure. Model evaluation metrics at different thresholds. 
```{r confusion matrix}

library(SDMTools)

predDec<- extract(criticalMap, testCREMP)

matrix09SDMTools<-confusion.matrix(CREMPunique$Acropora_cervicornis,predDec, threshold = 0.09)
matrix27SDMTools<-confusion.matrix(CREMPunique$Acropora_cervicornis,predDec, threshold = 0.27)

matrix_09<-accuracy(CREMPunique$Acropora_cervicornis,predDec, threshold = .09)
matrix_27<-accuracy(CREMPunique$Acropora_cervicornis,predDec, threshold = .27)

thresholds<-seq(0.03,0.3,.01)
AUCs<-acc<-sens<-spec<-areas<-list()

for(i in 1:length(thresholds)){
  predDec<- extract(criticalMap, testCREMP)
  th<-thresholds[i]
  matrix_i<-accuracy(CREMPunique$Acropora_cervicornis,predDec, threshold = th)
  AUCs[i]<-matrix_i$AUC
  sens[i]<-matrix_i$sensitivity
  spec[i]<-matrix_i$specificity
  acc[i]<-matrix_i$prop.correct
  areas[i]<-area(criticalMap > thresholds[i])
}

#S5 Figure. Model evaluation metrics at different thresholds. 
png( "S5Figure.png",  width= 5,  height= 4,  units = "in",
  res   = 300,  pointsize = 8)
par(mar=c(5,3,3,5))
plot(thresholds,AUCs, type="l", ylim=c(0,1), lwd=3, xlab = "Thresholds", ylab="")
lines(thresholds,sens, col = "blue")
lines(thresholds,spec, col = "orange")
lines(thresholds,acc, col = "purple", lwd=2)
lines(thresholds,as.double(areas)/3000, col = "green", lwd=2, lty=2)
axis(side = 4, at=seq(0,1.2,.2), labels=seq(0,3000,450))

mtext(side = 4, line = 3, expression(paste("Habitable Area (k ", m^{2})~ ")"))
abline(v=.09, lty=2)
abline(v=.27, lty=2)

legend(x = .21,y=0.52, 
       legend = c("AUC", "Specificity", "Sensitivity", "Accuracy", "Habitable Area"), lwd=c(3,1,1,2,2), lty = c(1,1,1,1,2),
       col = c("black","orange","blue", "purple","green"))
box(which = "plot")
dev.off()
```
